---
title: "Analysis"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, echo=FALSE}

batting<-read.csv("../EDA_Data/Batting.csv",header=TRUE)
pitching<-read.csv("../EDA_Data/Pitching.csv",header=TRUE)
salaries<-read.csv("../EDA_Data/Salaries.csv",header=TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)

suppressMessages(library(tidyverse))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))

```

```{r}
batting2 = filter(batting, batting$yearID > 1961) %>% select(yearID,H,X2B,X3B,HR,AB,BB,HBP,SF) %>% na.omit() %>% mutate(TB = H+2*X2B+(3*X3B)+(4*HR)) %>% mutate(SLG = TB/AB) %>% na.omit() %>% mutate(OBP = (H + BB + HBP)/(AB + BB + SF + HBP)) %>% mutate(OPS = SLG + OBP) %>% mutate(steroidEra = (yearID>1983 & yearID<2003))

table = batting2 %>% group_by(yearID,steroidEra) %>% summarise(avgOPS = mean(OPS))

ggplot(table, aes(x =yearID, y = avgOPS)) + geom_point(aes(x =yearID, y = avgOPS ,col=steroidEra)) + geom_smooth(se=FALSE) + ggtitle("Average OPS Over Time") + xlab("Year") + ylab("Average OPS")
```

### We can see that average OPS increases steadily during the 60's and 70's, increasing moderately still through 80's, plateu around 2000,  then decreases steadily after. 

----

```{r}
pitching2 = filter(pitching, pitching$yearID > 1961)%>%
  select(yearID,ERA) %>% na.omit %>%
  mutate(steroidEra = (yearID>1983 & yearID<2003))
table2 = pitching2 %>% group_by(yearID,steroidEra) %>% summarise(avgERA = mean(ERA))
ggplot(table2, aes(x =yearID, y = avgERA)) + geom_point(aes(col = steroidEra)) + geom_smooth(se=FALSE) + ggtitle("Average ERA Over Time") + xlab("Year") + ylab("Average ERA")  

```

###We can see that average ERA decreases until the 1980's, then increases drasticly up till the early 2000's, then begins a steady decrease.

----

*__From the 60's up till the steriod era, we see an increase in average OPS and decrease in average ERA, signifying that preformance was increasing for both pitchers and hitters.__* 


*__During the steriod era, ERA increase when OPS did, and then both begin to decrease around the end of the steriod era.__*

*__It appears that steriods did give an advantage to batters over pithers during the steriod era.__*

----

```{r}
batsal = left_join(batting,salaries, by="playerID")
batsal2 = filter(batsal, batsal$yearID.x > 1985) %>% select(yearID.x,H,X2B,X3B,HR,AB,BB,HBP,SF,salary,teamID.x) %>% na.omit() %>% mutate(TB = H+2*X2B+(3*X3B)+(4*HR)) %>% mutate(SLG = TB/AB) %>% na.omit() %>% mutate(OBP = (H + BB + HBP)/(AB + BB + SF + HBP)) %>% mutate(OPS = SLG + OBP) %>% mutate(steroidEra = (yearID.x>1983 & yearID.x<2003))

ggplot(batsal2, aes(x =OPS, y = log(salary))) + geom_point(aes(x =OPS, y = log(salary), col = steroidEra)) + geom_smooth(se=FALSE)

```

###The bottom line shows that salaries were increasing quicker when OPS was increasing.


```{r, include=FALSE}
#str(master_file)
#Things to consider for for RQ's: strikeouts and era for pitchers pre, during and post(steriod use).....also obp and batting average for hitters pre, during and post.
#Steriod use staring around late 80's and dying out in 2003, yet continuing to see suspensions to this day for steriods.
```




